---
title: "Statistics 101 - Homework 1"
header-includes: \usepackage{float}
output: pdf_document
---

```{r, echo=FALSE}
# disable the formatting of numbers to avoid error message
knitr::knit_hooks$set(inline = as.character)
```

### Name: ____________________________  Class Number: __________

\bigskip

__1)__ About the Mode: indicate whether the following statements are True or False

a. The mode is the most frequent value in a distribution.
b. There cannot be more than one mode per distribution.
c. The mode is influenced by extreme values.
d. The mode's value can change by organizing the data into different categories.
e. The mode is only applicable to qualitative variables.

\bigskip

__2)__ About the Median: indicate whether the following statements are True or False

a. The median is the central value with 50% of the values larger than it and 50% smaller.
b. There can be more than one median per distribution.
c. The median is affected by extreme values.
d. The median's value is rather stable even when data are organized into different categories.
e. The median is applicable to both qualitative and quantitative variables.

\bigskip

__3)__ About the Average: indicate whether the following statements are True or False

a. The average is the sum of all entries divided by half the number of entries.
b. There is only one average per distribution.
c. The average is not influenced by extreme values.
d. The average is applicable to quantitative variables only.
e. The average is always greater than the median.

\bigskip

__4)__ True or False. If a distribution is skewed to the right,

a. the mean is less than the median
b. the mean and the median are equal
c. the mean is greater than the median

\bigskip


__5)__ The highway mileages of 13 compact cars are:
```{r, echo=FALSE, results='asis', message=FALSE}
library(xtable)
car_name <- c(
  'Aston Martin Vanquish', 'Audi TT Coupe',
  'BMW 325CI', 'BMW 330CI', 'BMW M3', 'Jaguar XK8', 'Jaguar XKR', 
  'Lexus SC 430', 'Mini Cooper', 'Mitsubishi Eclipse', 
  'Mitsubishi Spyder', 'Porsche Cabriolet', 'Porsche Turbo 911')

mileage <- c(19, 29, 27, 28, 23, 26, 23, 23, 32, 31, 29, 26, 22)

carsdf <- data.frame(model = car_name, mileage = mileage)
xtb <- xtable(carsdf)
print(xtb, comment = FALSE, table.placement = 'H', type = 'latex')
```

a. Calculate the median:
b. Calculate the mode:
c. Calculate the mean:
d. Calculate the range:
e. Calculate $Q_1$ (the 1st quartile)
f. Calculate $Q_3$ (the 3rd quartile)
g. Calculate the Interquartile Range (IQR):
h. Calculate the Variance Range (Var):
i. Calculate the Standard Deviation (SD):


\bigskip

```{r echo = FALSE}
# number of employees
clerks <- 5
juniors <- 2

# salaries of employees
clerk_salary <- 35000
junior_salary <- 80000
owner_salary <- 320000
salaries <- c(rep(clerk_salary, clerks), rep(junior_salary, juniors), owner_salary)

meanSalary <- mean(salaries)
medianSalary <- median(salaries)
```

__6)__ A small accounting firm pays each of its `r clerks` clerks $`r clerk_salary`, 
$`r juniors` junior accountants `r junior_salary` each, and the firm's owner 
 $`r owner_salary`.

a. What is the mean salary paid at this firm?
b. How many of the employees earn less than the mean?
c. What is the median salary?

\bigskip

```{r echo = FALSE}
owner_salary <- 455000
salaries <- c(rep(clerk_salary, clerks), rep(junior_salary, juniors), owner_salary)
mean_salary <- mean(salaries)
median_salary <- median(salaries)
```

__7)__ The firm in the previous question gives no raises to the clerks and junior accountants, while the owner's take increases to $`r owner_salary`.

a. How does this change affect the mean?
b. How does it affect the median?

