---
title: "HW: Flow Control Structures and Functions with R"
author: "Prof. Sanchez"
date: "April 17, 2015"
output: pdf_document
---

## Introduction

This assignment has two purposes:

a) to familiarize you with control flow structures in R 
b) to introduce you to writing functions in R


-----

## Arithmetic average with loops

R provides the function `mean()` that can be applied to calculate the arithmetic 
mean (i.e. average) of a numeric vector. Here's the formula of the average for a 
vector $x = (x_1, x_2, \dots, x_n)$
$$
\bar{x} = \frac{1}{n} \sum^{n}_{i = 1} x_i
$$

Use the three types of loops---`for, repeat, while`---to get the average of the 
numeric vector  `x <- 1:10`


```{r, echo=FALSE, eval=FALSE}
x <- 1:10
total <- length(x)
x_sum <- 0
for (i in seq_along(x)) {
  x_sum <- x_sum + x[i]
}
x_sum / total
```


```{r, echo=FALSE, eval=FALSE}
x <- 1:10
total <- length(x)
x_sum <- 0
i <- 1
repeat {
  x_sum <- x_sum + x[i]
  if (i == total) break
  i <- i + 1
}
x_sum / total
```


```{r, echo=FALSE, eval=FALSE}
x <- 1:10
total <- length(x)
x_sum <- 0
i <- 1
while (i <= total) {
  x_sum <- x_sum + x[i]
  i <- i + 1
}
x_sum / total
```


-----

## Area of a circle

For a given circle of radius $r$, the area $A$ is:
$$
A = \pi r^2
$$

Write a function `circle_area()` that calculates the area of a circle. This 
function must take one argument `radius`. Give `radius` a default value of 1.

```{r, echo=FALSE}
circle_area <- function(radius = 1) {
  pi * radius^2
}
```

For example:
```{r}
# default (radius 1)
circle_area()

# radius 3
circle_area(radius = 3)
```


## Area of a cylinder

For a given cylinder of radius $r$ and height $h$ the area $A$ is:
$$
A = 2 \pi r h + 2 \pi r^2
$$

Notice that the formula of the area of a cylinder includes the area of a circle: 
$\pi r^2$. Write a function `cyl_area()`, that calls `circle_area()`, to compute 
the area of a cylinder. This function must take two arguments: `radius` and `height`. 
Give both arguments a default value of 1.

```{r, echo=FALSE}
cyl_area <- function(radius = 1, height = 1) {
  lateral_area <- 2 * pi * radius * height
  base_areas <- 2 * circle_area(radius)
  lateral_area + base_areas
}
```

For instance:
```{r}
# default (radius 1, height 1)
cyl_area()

# radius 2, height 3
cyl_area(radius = 2, height = 3)
```


## Volume of a cylinder

For a given cylinder of radius $r$ and height $h$ the volume $V$ is:
$$
V = \pi r^2 h
$$

Write a function `cyl_volume()`, that calls `circle_area()`, to compute the 
volume of a cylinder. This function must take two arguments: `radius` and `height`. 
Give both arguments a default value of 1.

```{r, echo=FALSE}
cyl_volume <- function(radius = 1, height = 1) {
  circle_area(radius) * height
}
```

For example:
```{r}
# default (radius 1, height 1)
cyl_volume()

cyl_volume(radius = 3, height = 10)
cyl_volume(height = 10, radius = 3)
```


-----

## Even number

Write a function `is_even()` that determines whether a number is even 
(i.e. multiple of 2). If the inpute number is even, the output should be `TRUE`. 
If the input number is odd, the output should be `FALSE`. If the input is not a 
number, the output should be `NA` 

```{r, echo = FALSE}
is_even <- function(x) {
  if (is.numeric(x)) {
    return(x %% 2 == 0)
  } else {
    return(NA)
  }
}
```

For example:
```{r}
# even number
is_even(10)

# odd number
is_even(33)

# not a number
is_even('a')
```


## Odd number

Use your function `is_even()` to write a function `is_odd()` that determines 
if a number is odd (i.e. not a multiple of 2). If a number is odd, the output 
should be `TRUE`; if a number is even the output should be `FALSE`; if the 
input is not a number the output should be `NA`

```{r, echo = FALSE}
is_odd <- function(x) {
  !is_even(x)
}
```

For example:
```{r}
# odd number
is_odd(1)

# even number
is_odd(4)

# not a number
is_odd('a')
```


-----

## Converting Miles to other units

The table below shows the different formulas for converting miles (mi) into other scales:

| Units   | Formula         |
| ------- | --------------- |
| Inches  | mi x 63360      |
| Feet    | mi x 5280       |
| Yards   | mi x 1760       |
| Meters  | mi / 0.00062137 |
| Kms     | mi / 0.62137    |

Write the following five functions for each type of conversion. Each function 
must take one argument `x` with default value: `x = 1`.

- `miles2inches()`
- `miles2feet()`
- `miles2yards()`
- `miles2meters()`
- `miles2kms()`

```{r, echo = FALSE}
miles2inches <- function(x = 1) {
  x * 63360
}

miles2feet <- function(x = 1) {
  x * 5280
}

miles2yards <- function(x = 1) {
  x * 1760
}

miles2meters <- function(x = 1) {
  x / 0.00062137
}

miles2kms <- function(x = 1) {
  x / 0.62137
}
```

For example: 
```{r}
miles2inches(2)
miles2feet(2)
miles2yards(2)
miles2meters(2)
miles2kms(2)
```
